---
title: "R Notebook"
output:
  md_notebook: default
  html_notebook: default
---

## Load data and package


```{r}
rm(list = ls())
source('script_loaddata.R')

library(sem)
library(polycor)
```

### printing covariance matrix for SEM

```{r}
subdata <- data[c('pc1', 'Area','Noon_Shade', 'Distance_to_Nearest','Aall', 'Pall', 'Hall')]
covdata <- cov(subdata)

```


### covariance pattern between the three species

```{r}
spp <- data[c('Aall', 'Pall', 'Hall')]
cor(spp)
```


### Alternative SEM 1: reverse the H --> P link

The reversed link P --> H was not significant (p = 0.099).The overall model would still pass (p = 0.16) but the p value was lower compare to the model with H --> P link.

```{r}
model0E <- specifyModel(file="./model/mod0E.txt", endog.variances=TRUE)
semHerb <- sem(model0E, covdata, N=38, data)

summary(semHerb)
```


## Alternative SEM 2: no species interaction

Model with no species interaction term would still pass to be acceptable (p = 0.12).
But AIC and Chi-square both suggest the model with (H --> P) link is prefered. 

```{r}
model0F <- specifyModel(file="./model/mod0F.txt", endog.variances=TRUE)
semHerb <- sem(model0F, covdata, N=38, data)

summary(semHerb)
```
