---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
## Load data and package


```{r}
rm(list = ls())
source('script_loaddata.R')

library(sem)
library(polycor)
```

#### Preparing covariance matrix for SEM

* Aall: mean population of _Uroleucon_ (aphid)
* Pall: mean population of _Paria_ (beetle)
* Hall: mean population of _Hesperotettix_ (grasshopper)

```{r}
cordata <- cor(data)
```

## Correlation pattern between the three species

_Uroleucon_ negatively correlates with _Paria_. From the SEM, we see the negative correlation was most likely the result of their different response to shading level.

_Paria_ negatively correlates with _Hesperotettix_. Comparing the selected SEM with the alternative SEM 1 below, we can infer that this is partially caused by the negative effect of _Hesperotettix_ on _Paria_. (Rather than the negative effect of _Paria_ on _Hesperotettix_.)

```{r}
spp <- data[c('Aall', 'Pall', 'Hall')]
cor(spp)
```



## SEM result: 

The original model output for the model presented in **Figure 4**.

This model is selected among other models for lowest AIC.

```{r}
model0 <- specifyModel(file="./model/mod0.txt", endog.variances=TRUE)
semHerb <- sem(model0, cordata, N=38, data)

summary(semHerb)
```

## Trim model0: remove links with p > 0.1

```{r}
model0A <- specifyModel(file="./model/mod0A.txt", endog.variances=TRUE)
semHerb <- sem(model0A, cordata, N=38, data)

S <- summary(semHerb)
S 
AIC0 <- S$AIC

```

## Add species interaction terms
1. Add one link each time

### Uroleucon -> Paria
```{r}
model1_UP <- specifyModel(file="./model/mod1_UP.txt", endog.variances=TRUE)
semHerb <- sem(model1_UP, cordata, N=38, data)
S <- summary(semHerb)

paste('Uroleucon -> Paria (',
      as.character(round(S$coeff$Estimate[1], digits = 2)), ')')
paste('AIC :', as.character(round(S$AIC, digits = 2)), 
      '(change:', as.character(round(S$AIC - AIC0, digits = 2)), ')')
paste('p =', as.character(round(S$coeff$`Pr(>|z|)`[1], digits = 2)))
```

### Paria -> Uroleucon 

```{r}
model1_PU <- specifyModel(file="./model/mod1_PU.txt", endog.variances=TRUE)
semHerb <- sem(model1_PU, cordata, N=38, data)
S <- summary(semHerb)

paste('Paria -> Uroleucon (',
      as.character(round(S$coeff$Estimate[1], digits = 2)), ')')
paste('AIC :', as.character(round(S$AIC, digits = 2)), 
      '(change:', as.character(round(S$AIC - AIC0, digits = 2)), ')')
paste('p =', as.character(round(S$coeff$`Pr(>|z|)`[1], digits = 2)))
```

### Uroleucon -> Hesperotettix
```{r}
model1_UH <- specifyModel(file="./model/mod1_UH.txt", endog.variances=TRUE)
semHerb <- sem(model1_UH, cordata, N=38, data)
S <- summary(semHerb)

paste('Uroleucon -> Hesperotettix (',
      as.character(round(S$coeff$Estimate[1], digits = 2)), ')')
paste('AIC :', as.character(round(S$AIC, digits = 2)), 
      '(change:', as.character(round(S$AIC - AIC0, digits = 2)), ')')
paste('p =', as.character(round(S$coeff$`Pr(>|z|)`[1], digits = 2)))
```

### Hesperotettix -> Uroleucon 
```{r}
model1_HU <- specifyModel(file="./model/mod1_HU.txt", endog.variances=TRUE)
semHerb <- sem(model1_HU, cordata, N=38, data)
S <- summary(semHerb)

paste('Hesperotettix -> Uroleucon (',
      as.character(round(S$coeff$Estimate[1], digits = 2)), ')')
paste('AIC :', as.character(round(S$AIC, digits = 2)), 
      '(change:', as.character(round(S$AIC - AIC0, digits = 2)), ')')
paste('p =', as.character(round(S$coeff$`Pr(>|z|)`[1], digits = 2)))
```

### Paria -> Hesperotettix
```{r}
model1_PH <- specifyModel(file="./model/mod1_PH.txt", endog.variances=TRUE)
semHerb <- sem(model1_PH, cordata, N=38, data)
S <- summary(semHerb)

paste('Paria -> Hesperotettix (',
      as.character(round(S$coeff$Estimate[1], digits = 2)), ')')
paste('AIC :', as.character(round(S$AIC, digits = 2)), 
      '(change:', as.character(round(S$AIC - AIC0, digits = 2)), ')')
paste('p =', as.character(round(S$coeff$`Pr(>|z|)`[1], digits = 2)))
```

### Hesperotettix -> Paria 
```{r}
model1_HP <- specifyModel(file="./model/mod1_HP.txt", endog.variances=TRUE)
semHerb <- sem(model1_HP, cordata, N=38, data)
S <- summary(semHerb)

paste('Hesperotettix -> Paria  (',
      as.character(round(S$coeff$Estimate[1], digits = 2)), ')')
paste('AIC :', as.character(round(S$AIC, digits = 2)), 
      '(change:', as.character(round(S$AIC - AIC0, digits = 2)), ')')
paste('p =', as.character(round(S$coeff$`Pr(>|z|)`[1], digits = 2)))
```

## Final model

```{r}
S
```
