---
title: "R Notebook"
output:
  md_notebook: default
  html_notebook: default
---

## Load data and package


```{r}
rm(list = ls())
source('script_loaddata.R')

library(sem)
library(polycor)
```

#### Preparing covariance matrix for SEM

* Aall: mean population of _Uroleucon_ (aphid)
* Pall: mean population of _Paria_ (beetle)
* Hall: mean population of _Hesperotettix_ (grasshopper)

```{r}
subdata <- data[c('pc1', 'Area','Noon_Shade', 'Distance_to_Nearest','Aall', 'Pall', 'Hall')]
covdata <- cov(subdata)
```




## SEM result: 

The original model output for the model presented in **Figure 4**.

This model is selected among other models for lowest AIC.

```{r}
model0D <- specifyModel(file="./model/mod0D.txt", endog.variances=TRUE)
semHerb <- sem(model0D, covdata, N=38, data)

summary(semHerb)
```

## Correlation pattern between the three species

_Uroleucon_ negatively correlates with _Paria_. From the SEM, we see the negative correlation was most likely the result of their different response to shading level.

_Paria_ negatively correlates with _Hesperotettix_. Comparing the selected SEM with the alternative SEM 1 below, we can infer that this is partially caused by the negative effect of _Hesperotettix_ on _Paria_. (Rather than the negative effect of _Paria_ on _Hesperotettix_.)

```{r}
spp <- data[c('Aall', 'Pall', 'Hall')]
cor(spp)
```


## Alternative SEMs

### Alternative SEM 1: reverse the H --> P link

The reversed link P --> H was not significant (p = 0.099).

The overall model would still pass (p = 0.16) but the p value was lower compare to the model with H --> P link.

```{r}
model0E <- specifyModel(file="./model/mod0E.txt", endog.variances=TRUE)
semHerb <- sem(model0E, covdata, N=38, data)

summary(semHerb)
```


### Alternative SEM 2: no species interaction

Model with no species interaction term would still pass to be acceptable (p = 0.12).

But AIC and Chi-square both suggest the model with (H --> P) link is prefered. 

```{r}
model0F <- specifyModel(file="./model/mod0F.txt", endog.variances=TRUE)
semHerb <- sem(model0F, covdata, N=38, data)

summary(semHerb)
```


### Alternative SEM 3: force a A --> H link

The  A --> H link (effect of _Uroleucon_ on _Hesperotettix_) was not significant (p =0.42) and was even negative.

```{r}
model0G <- specifyModel(file="./model/mod0G.txt", endog.variances=TRUE)
semHerb <- sem(model0G, covdata, N=38, data)

summary(semHerb)
```
